---
title: "RR-PA1"
author: "Mike Danielson"
date: "September 19, 2015"
output: html_document
---
Step-1: set-up environment, download & unzip file, import as data frame
```{r}
library(plyr)
library(lattice)
setInternet2(TRUE)
download.file(url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",destfile = "temp")
unzip("temp","activity.csv")
data <- read.csv("activity.csv")
```
Step-2: process data for further use
```{r}
hour <- trunc(data$interval/100)
data$datetime <- paste(data$date,":",hour,":",data$interval - 100*hour,sep="")
data$datetime <- strptime(data$datetime,"%Y-%m-%d:%H:%M")
```
Step-3: calc mean, median, and produce histogram
```{r}
avedata <- na.omit(aggregate(data$steps, by =list(as.Date(data$datetime)),FUN = mean,na.action=na.omit))
avedata <- rename(avedata,c("Group.1"="date","x"="steps"))
format(avedata,digits=4)
hist(avedata$steps,main ="Histogram of Steps/Day",xlab="Steps",ylab="Frequency")
paste("mean of steps per day =",format(mean(avedata$steps),digits=6))
paste("median of steps per day =",format(median(avedata$steps),digits=6))
```
Step-4: Create time series plot of average number of steps taken, averaged across all days.
Determine interval with maximum activity.
```{r}
avedata <- na.omit(data)
avedata <- aggregate(avedata$steps, by =list(avedata$interval),FUN = mean)
avedata <- rename(avedata,c("Group.1"="interval","x"="steps"))
plot(x = avedata$interval,y = avedata$steps,type="l",main ="Average Steps Taken",xlab="5-minute Interval",ylab="Steps")
paste("Maximum activity occurs during interval",data$interval[which.max(avedata$steps)])
```
Step-5: Calc number of NAs. Strategy for filling NAs and fill NAs in new dataset. I replaced the NAs with the mean of all
intervals for all days. Compare to original dataset.Determine interval with maximum activity.
```{r}
paste("total number of missing data values =",sum(is.na(data$steps)))
newdata <- data
newdata[is.na(newdata)] <- mean(newdata$steps,na.rm=TRUE)
avenewdata <- na.omit(aggregate(newdata$steps, by =list(as.Date(newdata$datetime)),FUN = mean,na.action=na.omit))
avenewdata <- rename(avenewdata,c("Group.1"="date","x"="steps"))
format(avenewdata,digits=4)
hist(avenewdata$steps,main ="Histogram of Steps/Day",xlab="Steps",ylab="Frequency")
paste("mean of steps per day =",format(mean(avenewdata$steps),digits=6))
paste("median of steps per day =",format(median(avenewdata$steps),digits=6))
```
Step-6: Create new factor variable for weekdays/weekends. Make a panel plot of weekend/weekday 5-minute data.
```{r}
newdata$weekend <- weekdays(newdata$datetime)=="Saturday" | weekdays(newdata$datetime)=="Sunday"
newdata$weekend <- factor(newdata$weekend, labels = c("weekday","weekend"))
avedata <- aggregate(newdata$steps, by =list(newdata$interval,newdata$weekend),FUN = mean)
avedata <- rename(avedata,c("Group.1"="interval","Group.2"="weekend","x"="steps"))
xyplot(avedata$steps~avedata$interval|avedata$weekend,type='l',layout=(c(1,2)),xlab="Interval",ylab="Number of steps")
```
